An efficient acyclic contact planner for multiped robots

### 2017-01-09 body/appendix_roms.tex
final version !
replaced contact feasible with contact reachable to match with reachability confition

\section{Generating the $W$ volumes for HRP-2}
\label{app:rom}

@@

Figure~\ref{fig:hrp2_roms} illustrate the obtained $W^k$ for HRP-2.
Regarding the procedure, we can see that step 2 is conservative (Figure~\ref{fig:hrp2_roms}--right), which 
is acceptable, especially because the lost set essentially relates to configurations close to singularity (they are close to the boundaries of the reachable workspace). We choose again to be less complete
but more efficient, regarding the number of collision tests to be performed by RB-RRT.
In step 1 on the other hand, selecting the convex hull (Figure~\ref{fig:hrp2_roms}--middle) instead of a minimum encompassing shape (Figure~\ref{fig:hrp2_roms}--left) may introduce false positives.
Concretely, because the false positive set intersects with $W^0$, the scaling volume of the robot torso, the induced error is compensated,
as verified by the results shown by Table~\ref{tab:requestpercent}.

>>>>

\clearpage
\section{Generating the $W$ volumes for HRP-2}
\label{app:rom}

@@

Figure~\ref{fig:hrp2_roms} illustrate the obtained $W^k$ for HRP-2.
Regarding the procedure, we can see that step 2 is conservative (Figure~\ref{fig:hrp2_roms}--right), which 
is acceptable, especially because the lost set essentially relates to configurations close to singularity (they are close to the boundaries of the reachable workspace, and
often not \gls{contact reachable}, as illustrated in Figure~\ref{fig:dedefeas}, where the exterior boundaries of the reachable workspace appear
red, thus not belonging to \gls{ccontact0}). We choose again to be less complete but more efficient, regarding the number of collision tests to be performed by RB-RRT.
In step 1 on the other hand, selecting the convex hull (Figure~\ref{fig:hrp2_roms}--middle) instead of a minimum encompassing shape (Figure~\ref{fig:hrp2_roms}--left) may introduce false positives.
Concretely, because the false positive set intersects with $W^0$, the scaling volume of the robot torso, the induced error is compensated,
as verified by the results shown by Table~\ref{tab:requestpercent}.

### 2017-01-09 body/glossary.tex
replaced contact feasible with contact reachable to match with reachability confition

\newglossaryentry{ccontact0}
{
  name=$C_{Contact}^0$,
  description={Set of \textit{contact feasible} root configurations $\mathbf{q}^{0}$, that verify $\exists \mathbf{q}^{\overline{0}},  \mathbf{q}^{0} \oplus \mathbf{q}^{\overline{0}} \in C_{contact}$}
}

\newglossaryentry{contact feasible}
{
  name=\textit{contact feasible},
  description={Refers to a root configuration that can be extended into a collision free contact configuration}
}

>>>>

\newglossaryentry{ccontact0}
{
  name=$C_{Contact}^0$,
  description={Set of \textit{contact reachable} root configurations $\mathbf{q}^{0}$, that verify $\exists \mathbf{q}^{\overline{0}},  \mathbf{q}^{0} \oplus \mathbf{q}^{\overline{0}} \in C_{contact}$}
}

\newglossaryentry{contact reachable}
{
  name=\textit{contact reachable},
  description={Refers to a root configuration that can be extended into a collision free contact configuration}
}

### 2017-01-09 body/intro.tex
replaced contact feasible with contact reachable to match with reachability confition

\noindent \textbf{Regarding $\mathcal{P}_1$}, we propose a fast guide path planning algorithm. The key to its efficiency is that it does not sample
directly the contact manifold, but an approximation of the \textit{contact feasible} space. The \textit{contact feasible} space is a low dimensional space for which there exists a mapping to the contact manifold. The space is sampled thanks to the \textit{reachability condition}, a low-dimensional geometric condition for \textit{contact feasibility}.

\noindent \textbf{Regarding $\mathcal{P}_2$},  we propose a fast method to extend a \textit{contact feasible} path into a sequence of whole-body configurations in static equilibrium. This  requires the explicit computation of contact configurations. It is guided by dedicated heuristics that quickly synthesize feasible configurations.

@@

\deladp{As a result, we propose contributions to both problems \Pa and $\mathcal{P}_2$.}
\begin{itemize}
\item \deladp{A low-dimensional, efficient sampling-based approach to plan guide paths.}
\item \deladp{A very efficient and general implementation of an acyclic contact planner, the first one compatible with \gls{interactive} applications.}
%~ \item Four heuristics for contact generation. They bias the planner towards configurations that are in robust static equilibrium, or 
%~ more efficient with respect to the task.
\end{itemize}

Finally, to demonstrate the validity of our approach, we show that the generated contact plans  can be successfully executed (problem  $\mathcal{P}_3$), either in simulation or on the real HRP-2 robot.

>>>>

\noindent \textbf{Regarding $\mathcal{P}_1$}, we propose a fast guide path planning algorithm. The key to its efficiency is that it does not sample
directly the contact manifold, but an approximation of the \textit{contact reachable} space. The \textit{contact reachable} space is a low dimensional space for which there exists a mapping to the contact manifold. The space is sampled thanks to the \textit{reachability condition}, a low-dimensional geometric condition for \textit{contact reachability}.

\noindent \textbf{Regarding $\mathcal{P}_2$},  we propose a fast method to extend a \textit{contact reachable} path into a sequence of whole-body configurations in static equilibrium. This  requires the explicit computation of contact configurations. It is guided by dedicated heuristics that quickly synthesize feasible configurations.

@@

Finally, to demonstrate the validity of our approach, we show that the generated contact plans  can be successfully executed (problem  $\mathcal{P}_3$), either in simulation or on the real HRP-2 robot. For HRP-2, we detail the complete computation times we obtained to address sequentially the three problems, and compare them with respect to related work, thus clearly demonstrating that our method is order of magnitude faster than any existing method.

### 2017-01-09 body/overview.tex
replaced contact feasible with contact reachable to match with reachability confition

%~ \begin{table*}
%~ \centering
%~ \begin{tabular}{ c | >{\arraybackslash}m{200pt} | c}
   %~ Robot configuration set &  \centering Definition: $\mathbf{q} \in$ set $\Rightarrow$?  & Root configuration set\\
 %~ \hline
 %~ \hline
   %~ $C_{contact}$ & $\mathbf{q}$ collision free; at least one end effector in contact. &  $\mathbf{q}^{0} \in C_{contact}^0$ is ``contact feasible''.  \\
 %~ \hline
   %~ $C_{Equil} \subset C_{contact} $ & $\mathbf{q}$  in static equilibrium &  $\mathbf{q}^{0} \in C_{Equil}^0$ is ``equilibrium feasible''. \\
 %~ \end{tabular}
%~ \caption{TODO.}
%~ \label{tab:glossary}
%~ \quad
%~ \end{table*}

Figure~\ref{fig:framework} illustrates our workflow.
\Pa and \Pb are addressed in a sequential fashion: we first plan a root guide path with the \textit{reachability condition}, before
extending it into a sequence of configurations in static equilibrium.
In this Section we describe this workflow and give some notations used throughout the paper. Some key concepts are introduced,
whose definitions can be found in the glossary at the end of this paper.

@@

\subsection{Computation of a guide path --- \Pa}
We first consider the problem of planning a root guide path. 
%The objective is to compute a path of root configurations that will allow for equilibrium. 
%A desired requirement is to preserve the completeness of the planner: it should be able to explore any valid guide path, but at the same time, any computed guide path must lead to a sequence of configurations in static equilibrium.
\begin{figure}[t]
\centering
  \begin{overpic}[width=1\linewidth]{figures/2D_feas}
		\put (15,) {$C_{Equil}^0$      $\subset$} 
		\put (47,) {$C_{Contact}^0$ $\approx$ } 
		\put (76,) {$C_{Reach}^0$} 
		%~ \put (5,27) {3.b)} 
		%~ \put (37,27) {4.a)} 
		%~ \put (68,27) {4.b)} 
	\end{overpic}
\caption{Illustration of several root configurations sets used in this paper in a 2D scene. Obstacles are violet, and units are in meters. To show the sets in a 2D representation, all the rotational joints of HRP-2 are locked in the shown configuration, such that a torso configuration
is only described by two positional parameters (x and y). The root of the robot is indicated with a black cross. To compute the reachable workspace, the point on the ankle indicated by a green cross was used. $C_{Equil}^0$ is included in $C_{Contact}^0$. $C_{Reach}^0$ approximates $C_{Contact}^0$. Depending on a parametrization, we can obtain $C_{Contact}^0 \subset C_{Reach}^0$. Considering the configurations around the top obstacle, we can observe a dramatic
divergence between  $C_{Equil}^0$  and $C_{Contact}^0$ when the problem is not \textit{\gls{cluttered}}.}
		   \label{fig:dedefeas}
\end{figure}
%
Ideally we would like to sample root configurations that are \gls{equilibrium feasible}. A root configuration is \gls{equilibrium feasible} if and only if there exists at least one joint configuration such that:
\begin{description}
\item[C1]: the robot is not in collision with the environment, 
%~ \item[C2] has at least one end-effector in contact with the environment, 
\item[C2]: the robot is in static equilibrium. 
\end{description}
Verifying C2 is computationally expensive because it requires the projection of the whole-body configuration into a contact posture, as well as the resolution of a linear program to compute the contact forces~\citep{Prete2016}.
For this reason, we are interested in finding an approximation of the \glslink{equilibrium feasible}{\textit{equilibrium feasibility}} that is computationally efficient.
A requirement for C2 is the condition:
\begin{description}
\item[C3]: at least one end-effector is in contact with the environment.
\end{description}
C1 and C3 define \glslink{contact feasible}{\textit{contact feasibility}}, which is easier to verify than \glslink{equilibrium feasible}{\textit{equilibrium feasibility}} because it does not require the computation of contact forces.
We make the assumption that for the \textit{\gls{cluttered} contact planning problem}, most of the times \gls{contact feasible} root configurations are also \gls{equilibrium feasible}.
For this reason we neglect C2 during the resolution of $\mathcal{P}_1$, and we consider it only when solving \Pb.
However, even verifying \glslink{contact feasible}{\textit{contact feasibility}} is computationally too expensive because it requires the projection of the whole-body configuration into a contact posture.

An intuitive description of \gls{contact feasible} configurations is ``close, but not too close'': close, because a contact surface must be partially included in the reachable workspace of the robot (represented for the right leg in Figure~\ref{fig:contact_gen}--a); not too close, because the robot must avoid collision.
We approximate \glslink{contact feasible}{\textit{contact feasibility}} with the \textit{reachability condition}, a geometrical criterion based on this intuitive description.
More precisely, a root configuration is \textit{reachable} if the root scaled by a factor $s \geq 1$ is not in collision, while the reachable workspace is in collision with the environment.
Verifying the \textit{reachability condition} is very efficient in practice.
In Section~\ref{rbprm} we detail how  we define and use the \textit{reachability condition} to compute a guide path with the Reachability-Based RRT (Figure~\ref{fig:framework}---\Pa).

Figure~\ref{fig:dedefeas} gives an insight into the difference between the three conditions, depicting $C_{Equil}^0$, $C_{Contact}^0$ and $C_{Reach}^0$, which are the sets of \gls{equilibrium feasible}, \gls{contact feasible} and \textit{reachable} root configurations, respectively.

In the remainder of this paper, we use the terms \gls{contact feasible} and \gls{equilibrium feasible} to qualify either a root configuration, a whole-body configuration, or a set
of such configurations.

>>>>

Figure~\ref{fig:framework} illustrates our workflow.
\Pa and \Pb are addressed in a sequential fashion: we first plan a root guide path with the \textit{reachability condition}, before
extending it into a sequence of configurations in static equilibrium.
In this Section we describe this workflow and give some notations used throughout the paper. 
\delst{Some key concepts are introduced,
whose definitions can be found in the glossary at the end of this paper.}

@@

\subsection{Computation of a guide path --- \Pa}
We first consider the problem of planning a root guide path. All root configurations 
As done by \citeauthor{Bouyarmane2009}, the aim is that 
%The objective is to compute a path of root configurations that will allow for equilibrium. 
%A desired requirement is to preserve the completeness of the planner: it should be able to explore any valid guide path, but at the same time, any computed guide path must lead to a sequence of configurations in static equilibrium.
\begin{figure}[t]
\centering
  \begin{overpic}[width=1\linewidth]{figures/2D_feas}
		\put (15,) {$C_{Equil}^0$      $\subset$} 
		\put (47,) {$C_{Contact}^0$ $\approx$ } 
		\put (76,) {$C_{Reach}^0$} 
		%~ \put (5,27) {3.b)} 
		%~ \put (37,27) {4.a)} 
		%~ \put (68,27) {4.b)} 
	\end{overpic}
\caption{Illustration of several root configurations sets used in this paper in a 2D scene. Obstacles are violet, and units are in meters. To show the sets in a 2D representation, all the rotational joints of HRP-2 are locked in the shown configuration, such that a torso configuration
is only described by two positional parameters (x and y). The root of the robot is indicated with a black cross. To compute the reachable workspace, the point on the ankle indicated by a green cross was used. $C_{Equil}^0$ is included in $C_{Contact}^0$. $C_{Reach}^0$ approximates $C_{Contact}^0$. Depending on a parametrization, we can obtain $C_{Contact}^0 \subset C_{Reach}^0$. Considering the configurations around the top obstacle, we can observe a dramatic
divergence between  $C_{Equil}^0$  and $C_{Contact}^0$ when the problem is not \textit{\gls{cluttered}}.}
		   \label{fig:dedefeas}
\end{figure}
%
Ideally we would like to sample root configurations that are \gls{equilibrium feasible}. A root configuration is \gls{equilibrium feasible} if and only if there exists at least one joint configuration such that:
\begin{description}
\item[C1]: the robot is not in collision with the environment, 
%~ \item[C2] has at least one end-effector in contact with the environment, 
\item[C2]: the robot is in static equilibrium. 
\end{description}
Verifying C2 is computationally expensive because it requires the projection of the whole-body configuration into a contact posture, as well as the resolution of a linear program to compute the contact forces~\citep{Prete2016}.
For this reason, we are interested in finding an approximation of the \glslink{equilibrium feasible}{\textit{equilibrium feasibility}} that is computationally efficient.
A requirement for C2 is the condition:
\begin{description}
\item[C3]: at least one end-effector is in contact with the environment.
\end{description}
C1 and C3 define \glslink{contact reachable}{\textit{contact reachability}}, which is easier to verify than \glslink{equilibrium feasible}{\textit{equilibrium feasibility}} because it does not require the computation of contact forces.
We make the assumption that for the \textit{\gls{cluttered} contact planning problem}, most of the times \gls{contact reachable} root configurations are also \gls{equilibrium feasible}.
For this reason we neglect C2 during the resolution of $\mathcal{P}_1$, and we consider it only when solving \Pb.
However, even verifying \glslink{contact reachable}{\textit{contact reachability}} is computationally too expensive because it requires the projection of the whole-body configuration into a contact posture.

An intuitive description of \gls{contact reachable} configurations is ``close, but not too close'': close, because a contact surface must be partially included in the reachable workspace of the robot (represented for the right leg in Figure~\ref{fig:contact_gen}--a); not too close, because the robot must avoid collision.
We approximate \glslink{contact reachable}{\textit{contact reachability}} with the \textit{reachability condition}, a geometrical criterion based on this intuitive description.
More precisely, a root configuration is \textit{reachable} if the root scaled by a factor $s \geq 1$ is not in collision, while the reachable workspace is in collision with the environment.
Verifying the \textit{reachability condition} is very efficient in practice.
In Section~\ref{rbprm} we detail how  we define and use the \textit{reachability condition} to compute a guide path with the Reachability-Based RRT (Figure~\ref{fig:framework}---\Pa).

Figure~\ref{fig:dedefeas} gives an insight into the difference between the three conditions, depicting $C_{Equil}^0$, $C_{Contact}^0$ and $C_{Reach}^0$, which are the sets of \gls{equilibrium feasible}, \gls{contact reachable} and \textit{reachable} root configurations, respectively.

In the remainder of this paper, we use the terms \gls{contact reachable} and \gls{equilibrium feasible} to qualify either a root configuration, a whole-body configuration, or a set
of such configurations.

### 2017-01-09 body/rbprm.tex
replaced contact feasible with contact reachable to match with reachability confition

Again, we assume that, for \gls{cluttered} problems, most of the times \contactfeasibility\, implies \equilibriumfeasibility.
Under this assumption, our goal is to find a \gls{contact feasible} guide path.

To generate such a path efficiently, ideally we need to only sample \gls{contact feasible} root configurations.
This requires exhibiting a necessary and sufficient condition for \contactfeasibility. %,that does not need the computation of the whole-body configuration.
By default, verifying \contactfeasibility\, implies a constructive demonstration by exhibiting a valid $\mathbf{q}^{\overline{0}}$. This is the approach chosen by \cite{Bouyarmane2009},
which is too computationally expensive.
To formulate a cheaper condition, we may turn our attention towards either an only-necessary or an only-sufficient condition.

Only-necessary conditions are appealing because they preserve the completeness of the search, while reducing 
the search space: they provide an outer approximation of \gls{ccontact0}.
On the other hand, only-sufficient conditions provide the guarantee that any configuration that satisfies them is indeed \gls{contact feasible}:
they provide an inner approximation of \gls{ccontact0}.

@@

\subsection{Conditions for contact feasibility}
%~ By default, \glslink{contact feasible}{\textit{contact feasibility}} implies a constructive demonstration by exhibiting a valid $\mathbf{q}^{\overline{0}}$. This is the approach chosen by \cite{Bouyarmane2009}. However, computing $\mathbf{q}^{\overline{0}}$ is expensive in terms of computation time. In this section we rather define a necessary condition and a sufficient condition for \glslink{contact feasible}{\textit{contact feasibility}} that do not require this explicit computation.

\subsubsection*{Contact feasibility, a necessary condition:}
For a contact to be possible, a volume $O_i \in O$ necessarily intersects with the reachable workspace $W(\mathbf{q}^{0})$ (Figure~\ref{fig:contact_gen}--1). Furthermore, if $\mathbf{q}^{0}$ is \gls{contact feasible}, then the torso of the robot $W^0(\mathbf{q}^{0})$ is necessarily not colliding  with the environment $O$.

@@

\subsubsection*{Contact feasibility, a sufficient condition:}
%~ In building $C_{reach}^1$, we directly considered the including hull of the root body $W^{0}$, obtaining a necessary condition. 
A trivial sufficient condition for \glslink{contact feasible}{\textit{contact feasibility}} can be constructed as a variation of $\mathcal{C}^0_{\textrm{\it Nec}}$, by replacing $W^0$ with a bounding volume $B^{\textrm{\it Suf}}$ encompassing the whole robot in a given pose, except for the effector surfaces to be in contact. We denote by \mbox{$\mathcal{C}^0_{\textrm{\it Suf}} \subset $ \gls{ccontact0}} the set of root configurations corresponding to this sufficient condition:

@@

\subsection{Contact feasibility: a compromise \gls{reachc}}
The sufficient condition is not interesting in practice since it leads the solver to lose too many interesting paths. The necessary condition is not perfect either, since the first stage of the planner would stop on a guide that is not \gls{contact feasible}. 
An ideal shape $B$ (with $W^0 \subset B \subset B^{\textrm{\it Suf}}$) that leads to a necessary and sufficient condition may exist---even if it seems intuitively very unlikely in general. 
%~ Its construction is out of the scope of this work.

@@

However, using a shape between $W^0$ and $B^{\textrm{\it Suf}}$ leads to a trade-off between a necessary and a sufficient condition. We define $W^0_s$ as the volume $W^0$ subject to a scaling transformation by a factor $s \in \mathbb{R}^+$.
%
We then consider the spaces $C_{s}^0$
 \begin{equation}
C^0_s = \{ \mathbf{q}^0 : W(\mathbf{q}^{0}) \cap O \neq \emptyset \text{ and } W^0_s(\mathbf{q}^{0}) \cap O = \emptyset \} % \\ }
 %~ & \text{ and } & A_{torso}(\mathbf{q}^{root}) \cap W = \emptyset \}
\end{equation}
%
If $s=1$, then $W^0_s$ = $W^0$, such that $C_1^0$ = $\mathcal{C}^0_{\textrm{\it Nec}}$. We thus consider that $s \geq 1$, since smaller values would only worsen the approximation.
By increasing $s$, the condition can become sufficient.
%~ The higher $s$, the closer this \gls{reachc} is to being sufficient, and if $s=1$, the search is complete (in the sense that all \gls{contact feasible} configurations can be found). 
The parametrization of $s$ then defines a trade-off between these two interesting extremes. 
We can choose $s$ by hand, or automatically as explained in Section~\ref{sec:params}.
The chosen value $s=s^*$ defines the \gls{reachc}, therefore we write \gls{creach0} $= C^0_{s^*}$.
%~ We call this condition the \gls{reachc}, which is defined by the chosen value $s=s^*$, and define \gls{creach0} $= C^0_{s^*}$.
%~ In the rest of the paper, we consider that an appropriate value $s^*$ of $s$ is chosen, and define \gls{creach0} $= C^0_{s^*}$.

@@

Thanks to these modifications, the problem of planning a \gls{contact feasible} path is reduced to a geometric collision-checking problem, of low dimension (6 for HyQ, 8 for HRP-2 that has 2 joints in the torso). By doing this we can solve the problem with a sampling-based approach, in \gls{interactive} computation times.

>>>>

Again, we assume that, for \gls{cluttered} problems, most of the times \contactreachability\, implies \equilibriumfeasibility.
Under this assumption, our goal is to find a \gls{contact reachable} guide path.

To generate such a path efficiently, ideally we need to only sample \gls{contact reachable} root configurations.
This requires exhibiting a necessary and sufficient condition for \contactreachability. %,that does not need the computation of the whole-body configuration.
By default, verifying \contactreachability\, implies a constructive demonstration by exhibiting a valid $\mathbf{q}^{\overline{0}}$. This is the approach chosen by \cite{Bouyarmane2009},
which is too computationally expensive.
To formulate a cheaper condition, we may turn our attention towards either an only-necessary or an only-sufficient condition.

Only-necessary conditions are appealing because they preserve the completeness of the search, while reducing 
the search space: they provide an outer approximation of \gls{ccontact0}.
On the other hand, only-sufficient conditions provide the guarantee that any configuration that satisfies them is indeed \gls{contact reachable}:
they provide an inner approximation of \gls{ccontact0}.

@@

\subsection{Conditions for contact reachability}
%~ By default, \glslink{contact reachable}{\textit{contact reachability}} implies a constructive demonstration by exhibiting a valid $\mathbf{q}^{\overline{0}}$. This is the approach chosen by \cite{Bouyarmane2009}. However, computing $\mathbf{q}^{\overline{0}}$ is expensive in terms of computation time. In this section we rather define a necessary condition and a sufficient condition for \glslink{contact reachable}{\textit{contact reachability}} that do not require this explicit computation.

\subsubsection*{contact reachability, a necessary condition:}
For a contact to be possible, a volume $O_i \in O$ necessarily intersects with the reachable workspace $W(\mathbf{q}^{0})$ (Figure~\ref{fig:contact_gen}--1). Furthermore, if $\mathbf{q}^{0}$ is \gls{contact reachable}, then the torso of the robot $W^0(\mathbf{q}^{0})$ is necessarily not colliding  with the environment $O$.

@@

\subsubsection*{contact reachability, a sufficient condition:}
%~ In building $C_{reach}^1$, we directly considered the including hull of the root body $W^{0}$, obtaining a necessary condition. 
A trivial sufficient condition for \glslink{contact reachable}{\textit{contact reachability}} can be constructed as a variation of $\mathcal{C}^0_{\textrm{\it Nec}}$, by replacing $W^0$ with a bounding volume $B^{\textrm{\it Suf}}$ encompassing the whole robot in a given pose, except for the effector surfaces to be in contact. We denote by \mbox{$\mathcal{C}^0_{\textrm{\it Suf}} \subset $ \gls{ccontact0}} the set of root configurations corresponding to this sufficient condition:

@@

\subsection{contact reachability: a compromise \gls{reachc}}
The sufficient condition is not interesting in practice since it leads the solver to lose too many interesting paths. The necessary condition is not perfect either, since the first stage of the planner would stop on a guide that is not \gls{contact reachable}. 
An ideal shape $B$ (with $W^0 \subset B \subset B^{\textrm{\it Suf}}$) that leads to a necessary and sufficient condition may exist---even if it seems intuitively very unlikely in general. 
%~ Its construction is out of the scope of this work.

@@

However, using a shape between $W^0$ and $B^{\textrm{\it Suf}}$ leads to a trade-off between a necessary and a sufficient condition. We define $W^0_s$ as the volume $W^0$ subject to a scaling transformation by a factor $s \in \mathbb{R}^+$.
%
We then consider the spaces $C_{s}^0$
 \begin{equation}
C^0_s = \{ \mathbf{q}^0 : W(\mathbf{q}^{0}) \cap O \neq \emptyset \text{ and } W^0_s(\mathbf{q}^{0}) \cap O = \emptyset \} % \\ }
 %~ & \text{ and } & A_{torso}(\mathbf{q}^{root}) \cap W = \emptyset \}
\end{equation}
%
If $s=1$, then $W^0_s$ = $W^0$, such that $C_1^0$ = $\mathcal{C}^0_{\textrm{\it Nec}}$. We thus consider that $s \geq 1$, since smaller values would only worsen the approximation.
By increasing $s$, the condition can become sufficient.
%~ The higher $s$, the closer this \gls{reachc} is to being sufficient, and if $s=1$, the search is complete (in the sense that all \gls{contact reachable} configurations can be found). 
The parametrization of $s$ then defines a trade-off between these two interesting extremes. 
We can choose $s$ by hand, or automatically as explained in Section~\ref{sec:params}.
The chosen value $s=s^*$ defines the \gls{reachc}, therefore we write \gls{creach0} $= C^0_{s^*}$.
%~ We call this condition the \gls{reachc}, which is defined by the chosen value $s=s^*$, and define \gls{creach0} $= C^0_{s^*}$.
%~ In the rest of the paper, we consider that an appropriate value $s^*$ of $s$ is chosen, and define \gls{creach0} $= C^0_{s^*}$.

@@

Thanks to these modifications, the problem of planning a \gls{contact reachable} path is reduced to a geometric collision-checking problem, of low dimension (6 for HyQ, 8 for HRP-2 that has 2 joints in the torso). By doing this we can solve the problem with a sampling-based approach, in \gls{interactive} computation times.
